---
title: "Descriptive Statistics with R"
author: "Qichang Feng"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Descriptive Statistics
for [Communities and Crime datest](https://archive.ics.uci.edu/ml/datasets/Communities+and+Crime)

#### Load stringr library to manipulate string
```{r} 
library(stringr)
```

#### Load the dataset
```{r}
dat <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/communities/communities.data", header = FALSE, stringsAsFactors = FALSE)
names <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/communities/communities.names", header = FALSE, skip = 75, nrows = 128, stringsAsFactors = FALSE)
dim(dat)
```

#### Extract the names for the attributes
```{r}
col_names <- str_trim(str_extract(names[, 1], "\\s\\w+\\s"))
colnames(dat) <- col_names
head(dat)
```

#### Descriptive Statistics for the dataset
```{r}
str(dat)
summary(dat)
```

#### Selecting varible containing no missing value
```{r}
selected <- colSums(dat == "?") == 0
datComplete <- dat[,selected]
dim(datComplete)
head(datComplete)
```

```{r}
CrimePerPop <- dat[, 128]
datNumeric <- apply(datComplete[, -2], 2, as.numeric)
```

#### Compute the correlation among the variables to the response
```{r} 
cor_vec <- cor(datNumeric, CrimePerPop)
```

#### Identify the top 9 most predictive variables with respect to response
```{r}
ix_ordered <- order(abs(cor_vec), decreasing = TRUE)
print("The top 9 most predictive variables are")
print(colnames(datNumeric)[ix_ordered[1:9]])
print(cor_vec[ix_ordered[1:9]])
```

#### Pair plots for these 9 varialbes
```{r}
pairs(datNumeric[, ix_ordered[1:9]])
```

#### Histogram for these variables
```{r}
par(mfcol=c(1, 2))
for (i in 1:9) {
  var_name <- colnames(datNumeric)[ix_ordered[i]]
  title_hist <- paste("Histogram of ", var_name, sep=" ")
  title_box <- paste("Boxplot of ", var_name, sep=" ")
  hist(datNumeric[, ix_ordered[i]], probability = TRUE, xlab = var_name, main = title_hist)
  lines(density(datNumeric[, ix_ordered[i]]), col="red", lwd=2)
  boxplot(datNumeric[, ix_ordered[i]], xlab = var_name, main = title_box)
}
```

```{r}
# corVector <- c()
# for (i in 1:ncol(datNumeric)) {
#   corVector <- c(corVector, cor(as.numeric(datNumeric[, i]), CrimePerPop))
# }
```
